---
title: "lab_02"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into the following options - under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus - and calculates how many loans are in each group. How would you describe the distribution of loans in a sentence?
Hint: You will need to use `case_when` to make this work.
**A1** The majority of the loans, 15,089, were between 10k and 50k, 4,150 loans were between 100k and 500k for the second largest loan range, and only 356 loans were 1 million +, making up the smallest amount range.

1 100k_500k           4150
2 10k_50k            15089
3 1m_plus              356
4 500k_1m              538
5 50k_100k            3976
6 under_10k           9227

```{r}
# case_when example
new_amount_grouping <- west_virginia_ppp %>%
  #in the step above, you're creating a new dataset or container called new_amount_grouping from the pre-existing west_virginia_ppp.
mutate(
  amount_by_grouping = case_when(
    #now you're mutating, or creating a new column called "amount_by_grouping" in the new container above, grouped by loan amount.
    amount < 10000 ~ 'under_10k',
    amount >= 10000 & amount < 50000 ~ '10k_50k',
    amount >= 50000 & amount < 100000 ~ '50k_100k',
    amount >= 100000 & amount < 500000 ~ '100k_500k',
    amount >= 500000 & amount < 1000000 ~ '500k_1m',
    amount >= 1000000 ~ '1m_plus'
    #now you're giving a range on the left, and telling what those ranges should be called on the right in green.
  )
)
# and then in the console, run the following line:
# new_amount_grouping %>% group_by(amount_by_grouping) %>% summarize(count =n()) #
```

**Q2** Some of the loans appear not to have been paid out - there is a column representing the amount of money that hasn't been disbursed to the applicant. Which servicing lenders have the most loans with undisbursed amounts greater than 0?
**A2** Davis Trust Company has the most undispursed loan amounts: 3 of $10,000, 2 of $7,000, 3 of $2,000 and 4 of $1,000 totaling $54,000 over 12 loans.

```{r}
west_virginia_ppp %>%
  select(lender, undisbursed_amount) %>%
arrange(desc(undisbursed_amount))
```

```{r}

new_undispursed_grouping <- west_virginia_ppp %>%
mutate(
  outstanding_undispursed = case_when(
undisbursed_amount == 10000  ~ '10k',
undisbursed_amount == 10000  ~ '10k',
undisbursed_amount == 2000  ~ '2k',
undisbursed_amount == 1000  ~ '1k',
undisbursed_amount >= 0 & undisbursed_amount <= 1000 ~ '0_to_1000',
  )
)
# and then in the console, run the following line:
# new_undispursed_grouping %>% group_by(undisbursed_amount) %>% summarize(count =n()) #

## I couldn't figure out how to get the names to also appear on the new grouping that I created in the console. I was able to see it in the data, create the new groupings based on the outstanding loan amounts, but I couldn't figure out a way to show the grouping of each loan amount with the lender names.
```

**Q3** What are some potential reasons why a loan might not be paid out? What questions does that circumstance raise for you?
**A3** 2 of the 3 $10,000 undisbursed amounts are for loans in the 100,000s. So I'd guess those loans are being paid in various installments overtime. That would make me inclined to look into the legitimacy of a multi-installation payment, looking at if there are any red flags that raise concern.


**Q4** Which month in 2021 had the most loans approved?
**A4** March

```{r}
west_virginia_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
```

**Q5** How many loans went to West Virginia applicants that listed the maximum number of jobs retained (500)? Which one had the earliest approved date?
**A5** 10, earliest of those was approved on 2020-04-08.

```{r}

new_wv_jobs_categories <- west_virginia_ppp %>%
mutate(
  wv_jobs_retained = case_when(
    jobs_retained == 500 ~ '500'
  )
)
# and then in the console run the following code:
    #new_wv_jobs_categories %>% group_by(wv_jobs_retained) %>% summarize(count =n())


# I tried to have this as a new command:
      #arrange(desc(date_approved))
    #not sure why this part of the code isn't working isn't working...
```
