---
title: "Class Reference"
author: "David Andrews"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this Dave's personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

To set the working directory, click on more, then set as working directory. Command as follows: setwd("~/Documents/GitHub/data_journalism_2022_spring")

To use the tidyverse, you need to "run" the library tidyverse. Need to do that in every notebook.

Can highlight a line and choose RUN in the tool bar above the notebook.

Run code like this:

```{r}
library(tidyverse)
```

click the Green arrow at the end of the line to run that block. "Run all the code in this current block."

```{r}
ppp_maryland_loans <- read_rds("pre_lab_01.rds")
```

^ Reads the data of the other prelab document in the folder, allows you to read it.

Gimpse gives you a look at a dataframe.

```{r}
head(ppp_maryland_loans)
```

--> shows you the top 6 rows of a document.

Console is like an exhaust pipe. The Console is the computer, the top is the Monitor. You can decide what you can see and hide.

When you're viewing the PPP Loans page, you can sort by each category. 


```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  )
```

This code is saying --> Start with PPP loans --> "AND THEN" (%>%), group by each county in Maryland. "AND THEN" summarize by counting up the number of loans in each county bucket.

"Give me the N of that", that means "count up the number of things."


```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))
```

--> ^ Same code as before, but once you've counted them up, arrange them in descending order from highest to lowest."


Summarizing data is one of the best thing to learn what's wrong with it. If the highest race is "unnamed" for every county application, we can't include race in a story about PPP loans. 

```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount)
  ) %>%
  arrange(desc(mean_loan_amount))
```

^^ Same, but then we're going to sum the totals.




Tuesday, February 1, & prelab 2 - 


Assigning a variable → 

Within the code block, put the variable name first before the document name, before the commands that bring you your result that you’re storing:

```{r}
wv_summary <- wv_ppp_loans %>%
  group_by(state) %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(desc(max_loan_amount))
```

How to arrange by ascending order:
arrange(header name)

So, arrange() by default puts them in ascending order.


If you run a block of code, and you get a block of red error text, and you run it again and it works, it was just a warning.

A "read" command can unzip zipped CSV files, R can read it. Really good if you're concerned about space.

Highlight a line of text --> command + Enter, runs the line of selected code. 


```{r}
head(maryland_ppp)
summary(maryland_ppp)
colnames(maryland_ppp)
glimpse(maryland_ppp)
```

^^ This gives you quick looks at a lot of the codes we ran previously.

Data types: dbl = a number.

```{r}
maryland_ppp %>%
  select(loan_number, amount, payroll_proceed) %>%
  mutate(
  percent_payroll = payroll_proceed/amount
)
```

^^mutate(new_column = value)
This means you're creating a new column that doesn't exist, and you're defining how the column functions. So when you say mutute -> name of new column equals the payroll proceed divided by the amount. 

You can also make sure everything is uppercase. --> 
#mutate(name = str_to_uppercase)

A double == sign means to test for a value.

Can run the following code in the console:
maryland_with_in_out %>% group_by(in_out) %>% summarize(count =n())

Lubridate - helps to parse dates and durations of time.

```{r}
# if/else example
maryland_with_in_out <- maryland_ppp %>%
  mutate(
    in_out = if_else(
        servicing_lender_state == 'MD', "IN", "OUT"
      )
  )
# if it is Maryland, its marked in. If its not Maryland, its marked out. 
```

Task 9: Create a new column using case_when
One common use of mutate is to create a new column and then populate it based on values in an existing column. For example, if we wanted to put the values in `jobs_retained` into broader categories in a new column called `jobs_retained_category` and we have multiple categories. We can use a function called `case_when` to make that easier.

**Task** Run the following code to create a new column called `jobs_retained_category` based on the values of `jobs_retained`. Add a description of what this code does to your reference notebook.

```{r}
# case_when example
maryland_jobs_categories <- maryland_ppp %>%
mutate(
  jobs_retained_category = case_when(
    jobs_retained < 10 ~ 'under_10',
    jobs_retained >= 10 & jobs_retained < 25 ~ '10_to_25',
    jobs_retained >= 25 & jobs_retained < 50 ~ '25_to_50',
    jobs_retained >= 50 & jobs_retained < 100 ~ '50_to_100',
    jobs_retained >= 100 & jobs_retained < 250 ~ '100_to_250',
    jobs_retained >= 250 & jobs_retained < 500 ~ '250_to_500',
    jobs_retained == 500 ~ '500'
  )
)
# and then in the console, run the following line:
# maryland_jobs_categories %>% group_by(jobs_retained_category) %>% summarize(count =n()) #
```

```{r}
prince_georges <- maryland_ppp %>% filter(project_county_name == "PRINCE GEORGES")
head(prince_georges)
```

^^ this will limit the data to the specific county name.

```{r}
maryland_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
```

^^ this allows you to see the data by what date it was approved. It then groups the loans per month and shows you which date had the most amount of loans approved.

-

Thursday, February 3 & lab 2 - 

You can log into R and click "run all" and it'll run all the code in the lab/prelab.

mutate(project_city = str_to_upper(project_city))
^ this make everything uppercase.

```{r}
and_prince_georges <- maryland_ppp %>% filter(project_county_name == "PRINCE GEORGES" & business_type == "Non-Profit Organization" & amount > 150000)

nrow(and_prince_georges)
```

^If that "&" between the non profit and amount is flipped to an “OR,” you’re getting wayyyyy more records because more gets accepted into the qualifications. 

IN the prelab 2, we had a lubridate command that was:
CleanDate = ymd_hms(date_approved). But we didn't have HMS (hour minute second) in the dataset, so we only end up running ymd(date_approved). That gives us a clear <date> field.

```{r}
maryland_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
# allows you to group by month, by the first of each month. Hence "floor date". There's also a ceiling date, which rounds it up to the next month, which is kinda bad. You might only use that if you're talking about a deadline, so you want to see everything that came in before that.
```


Pre-lab 3:

```{r}
dataset_name %>% 
mutate(id = as.character(id))
# This changes the id field, which is often interpreted as a number, to a character type instead.
# English translation: change=the id field from its current number type, make it appear as a character type (the number here will be the id number that appears in each row)
```

```{r}
ppp_maryland_loans %>%
  mutate(month_year_approved = floor_date(date_approved, "month")) %>%
  #create a new column called month_year_approved, which equals the floor date of the month of the date approved.
  group_by(month_year_approved) %>%
  #group the new month_year_approved column by...
   summarise(
    count=n()
  ) %>%
  #...summing their values..
  arrange(month_year_approved)
#... and arranging them by months
```

The "could not find function "%>%” message" almost always means you don’t have the tidyverse loaded. Load that library and try again.

If the GitHub commit is too large, either zip the dataset file before commiting or leave it out of the commit all together.

if you have to load something and it restarts the R session, you likely have to reload the library that you're working in. So the tidyverse, janitor, etc.

```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names()
#this changed the 1_id column name to x1_id, making it easier to clean up.

# display the cleaned dataset
cleaned_arnold_md_loans
```

```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id)
#this gives a chance to specifically rename all instances with numbers before _id and change them to id.

# display the cleaned dataset
cleaned_arnold_md_loans


```


```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct()
#this cuts down on duplicates, shows you the number of distinct rows. More accurate for everything.


# display the cleaned dataset
cleaned_arnold_md_loans

```

```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L))
#this makes all of the zip codes a uniform 5 numbers instead of the added on 4.

# display the cleaned dataset
cleaned_arnold_md_loans

```

```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  mutate(city = str_to_title(city))
#Standardizes the way a title (city) appears, like this: Baltimore


# display the cleaned dataset
cleaned_arnold_md_loans

```

```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  mutate(city = str_to_title(city)) %>%
  mutate(city = case_when(
    city == "Anold" ~ "Arnold",
    TRUE ~ city
    #turns misspellings into the correct name, and only does it if its within the city column.
  ))


# display the cleaned dataset
cleaned_arnold_md_loans

```

```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  mutate(city = str_to_title(city)) %>%
  mutate(city = case_when(
    city == "Anold" ~ "Arnold",
    #above, the exact match method.
    str_detect(city,"^Arnl") ~ "Arnold",
    TRUE ~ city
    #above, the standardize detect method.
  ))


# display the cleaned dataset
cleaned_arnold_md_loans

```
