---
title: "Reverse Engineering Project"
author: "Dave Andrews, Maddy Peek, Eve Sampson"
date: "3/5/22"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Nearly 250 women have been fatally shot by police since 2015](https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
fatal_police_shootings_data <- read_csv('data/fatal-police-shootings-data.csv')
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**:  “The Post began tracking fatal police shootings in a database in January 2015, months after a White police officer in Ferguson, Mo., killed Michael Brown, an unarmed Black man during a confrontation. Since then, police have shot and killed about 1,000 people a year.”
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?] Yes. Grouping the dates by month and then again by year, there have been about 1,000 people fatally shot by police every year since 2015.

```{r}
fatal_police_shootings_data_year <- fatal_police_shootings_data %>%
  mutate(Month = floor_date(date, "month")) %>%
  mutate(year = year(date)) #%>% 
 #group_by(year) %>%
 #summarize(
    #count = n()
 #)
```


### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here] - “About 31 percent, or 77, of the 247 women fatally shot by police since 2015 had mental health issues, compared to 22 percent of the 5,362 men killed.”
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?] We came pretty close. Instead of 77 of the 247 women noted in the sentence, we got 79/248 filtering for the conditions outlined in the sentence. We did, however, get very similar percentages, successfully illustrating the point of the sentence. We think there are a few unknowns that account for these slight variations in data. 1. We don't know if additional data was entered for this timeframe after the article was published. Perhaps more cases were added/more conditions were included that became known after the published date. 2. We don't know what specific date was used as the cutoff date when this article was published. Since it was published on 9/4/20, we figured any entries that occured on 9/4/20 itself was likely not included in the statistics. Playing around with the cutoff date as early as 9/1/20 didn't produce exact results either. Still, we were able to reproduce the overarching point of the sentence, which is that mental health was a resulting factor in about 31% of women fatally shot by police since 2015, while mental health only accounted for about 22% of the men fatally shot by police since 2015.

```{r}
# Put code to reverse engineer sentence here
fatal_police_shootings_before_article <- fatal_police_shootings_data_year %>%
  filter(date < "2020-09-04")
#this gets us all entries that happened before 2020-09-04 when the article was published.
```

```{r}
women_shot_mentalhealth <- fatal_police_shootings_before_article %>%
filter(gender == 'F') %>% 
#nrow() = 248 rows
filter(signs_of_mental_illness == 'TRUE') %>% 
  filter(manner_of_death == "shot") #%>% 
 # nrow() = 79
#(79/248)*100 = 31.8%

#79 of the 248 of women who were fatally shot by police, or 31.8%, were reported to show signs of mental illness.

# Display results of code below this codeblock
```

```{r}
men_shot_mentalhealth <- fatal_police_shootings_before_article %>%
filter(gender == 'M') %>% 
#nrow() = 5362
filter(signs_of_mental_illness == 'TRUE') %>% 
filter(manner_of_death == "shot") #%>% 
#nrow() = 1,153
#(1153/5362)*100 = 21.5%

#21.5% of men who were fatally shot by police were reported to show signs of mental illness.

# Display results of code below this codeblock
```


### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here] “Since 2015, police have fatally shot nearly 250 women. Like Taylor, 89 of them were killed at homes or residences where they sometimes stayed.”
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?] We were able to confirm the first part of the sentence; by filtering for gender and dates before publication, we got 248 fatalities. We are unable to discern if they were killed at their homes or residences where they stayed based on the data. We would need to check the locations against their home addresses to verify.

```{r}
# Put code to reverse engineer sentence here

women_shot_since_2015 <- fatal_police_shootings_before_article %>% 
  filter(gender == "F")
# %>% 
#nrow() = 248

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [“Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known. But within this small subset, Black women, who are 13 percent of the female population, account for 20 percent of the women shot and killed and 28 percent of the unarmed deaths”]

* **Analysis summary**: The findings have changed slightly as the data increased over time since publication, but the general sentiment of the findings in the article is the same. Through the data, we were able to confirm that, from 2015 - Mid 2020, Black women accounted for .85% of all fatal shootings in cases where race was known, or less than 1% as the article states. The data provided doesn't have population figures, so we took the pretense that Black women make up 13% of the female population as a fact. We were able to confirm that Black women account for 19.35% of the women shot and killed, as the sentence says 20%. We were also able to confirm that Black women account for ___ of the unarmed deaths, or 28% as the article states. Unfortunately, we were not able to confirm that Black women account for 28% of the total unarmed deaths. Instead, we got that they only account for 1.85%. 


```{r}
#Part one of the sentence - "Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known."
black_females_shot <- fatal_police_shootings_before_article %>% 
  filter(gender == "F" & race == "B")
#number of total police shootings= 5,611
#number of Black females shot= 48
(48/5611)*100
#Percentage of Black women that account for fatal police shootings: .85% (which matches the first sentence that Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known.)
```

```{r}
#Part two of the sentence - "But within this small subset, Black women, (who are 13 percent of the female population,) account for 20 percent of the women shot and killed...
women_shot_since_2015 <- fatal_police_shootings_before_article %>% 
  filter(gender == "F")
# women shot since 2015 = 248
48/248*100
#percentage of Black females shot (48) divided by the total number of women shot (248) (x100) is: 19.35%
```

```{r}
# Part three of the sentence - "...Black women...account for 28 percent of the unarmed deaths."
unarmed_shootings <- fatal_police_shootings_before_article %>%
  filter(armed == "unarmed")
#number of total unarmed shootings: 378
blackfem_unarmed_shootings <- unarmed_shootings %>% 
  filter(gender == "F" & race == "B")
7/378*100
#Percentage of Black, female, unarmed shootings (7) divided by the total number of unarmed shootings(378) (x100) is 1.85%

#1.85% does not equal 28%, as claimed in the sentence.
```


### Sentence 5

* **Sentence text**: [“The starkest difference between women and men is the rate: Women account for about half the population, but 4 percent of the killings. Of those fatally shot every year, about 44 have been women.”]
* **Analysis summary**: [According to the data prior to the article's publication, 4.4 percent of those killed by police have been women, like what the article claimed. Our analysis also showed that between 2015 and 2019, an average of 44.4 women were killed each year. Though the article does include some 2020 data, the year is not complete so it does not appear it was used to calculate this number. ]

```{r}
#women are 4 percent of all killings 
  women_percent <- fatal_police_shootings_before_article %>%
  filter(gender == "F") %>%
  summarize(
    count = n()
  )
#total number of women killed: 248
#total number of killings: 5,611

(248/5611)*100
# = 4.4%, confirming women account for 4% of police shootings.

#44 women killed per year - 
  women_total <- fatal_police_shootings_before_article %>%
    filter(gender == "F") %>%
      group_by(year) %>%
     summarize(
    count = n())
```

-30-