---
title: "lab_03"
author: "sean mussenden"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
install.packages("tidyverse")
library(tidyverse)

# Load janitor here
install.packages('janitor')
```

## Load data
```{r}
# Write code to load the WV slice of the ppp loan data here and call it wv_ppp_loans.
# The data is called lab_03.rds, and is in the same folder as lab_03.Rmd (this data notebook)
wv_ppp_loans <- read_rds("lab_03.rds")

```

## Answer questions

Q1: Which five cities located in Kanawha County, West Virginia, got the largest _number_ of loans? How many did each get? To answer this question, you'll have to clean the very messy "city" field.


Note: To answer this question, you'll need to see every variation of the "city" column in Kanawha County, and then fully clean the city field for those cities. Hint: once cleaned, the dataframe that counts the number of loans by city should have 38 distinct cities for Kanawha County.

A1: 1. Charleston (2802), 2. South Charleston (478), 3. Saint Albans (296), 4. Dunbar (162), 5. Nitro (140)

```{r}

new_cleaned_wv_cities <- wv_ppp_loans %>%
  mutate(city = str_to_title(city)) %>%
  mutate(city = case_when(
    city == "Balle" ~ "Belle",
    city == "Charelston" ~ "Charleston",
    city == "Crosslanes" ~ "Cross Lanes",
    city == "Elkveiw" ~ "Elkview",
    city == "So Charleston" ~ "South Charleston",
    city == "St Albans" ~ "Saint Albans",
    city == "St.albans" ~ "Saint Albans",
    city == "St. Albans" ~ "Saint Albans",
    city == "S Charleston" ~ "South Charleston",
    TRUE ~ city
  )
  ) %>% 

filter(project_county_name == "KANAWHA") %>%
group_by(city) %>%
  summarise(total = n()) %>%
  # first, arrange(city)
  arrange(desc(total)
  )

```

Q2: Find the 5-digit ZIP code in West Virginia that had the highest total amount of loans (by dollar value). What is the name of the business that got the largest loan (by dollar value) in that zip code? Write up a short paragraph that also answers these questions:

* Where in West Virginia is this ZIP code located?
* How much was the business' loan for?
* How much of that loan was forgiven?
* Which number is bigger, the loan amount or the forgiveness amount? Can you explain the discrepancy?

Notes: 
clean zip --> 5 
group by zipcode
arrange by desc(amount)


A2: Zip code 25301 had the most loans, totaling $160,721,321. JACKSON KELLY PLLC received the biggest loan in that zip code, totaling $5,068,472. It is located in Kanawha County, which encompasses the city of Charleston and congressional district WV-02. $5,126,618.6 of the loan was forgiven, which is more than the loan amount itself. I originally wondered if the difference was equal to the distance between the initial amount approved and the final amount paid, but those numbers are not the same. My best guess would be interest, but 50,000 in interest seems like too high of a value.

```{r}

new_wv_zips <- wv_ppp_loans %>%
  mutate(zip = str_sub(zip, start=1L, end=5L))
# Hint
# First, create a dataframe that shows the top 5-digit ZIP code by total dollar amount
# Second, use the top ZIP code to create a separate dataframe that shows the business with the biggest loan in that ZIP code.
top_zips <- new_wv_zips %>% 
  group_by(zip) %>% 
  summarise(total = sum(amount)) %>%
arrange(desc(total))
zip_25301_loans <- new_wv_zips %>%
  filter(zip == "25301") %>%
  arrange(desc(amount)
  )

```

Q3: Five business entities owned by the spouse of a member of West Virginia's U.S. House delegation got PPP loans. Write up a paragraph that answers the following questions.   

  * What are the businesses' names?
  * What do the businesses do?
  * How much money did the five businesses receive in total?
  * What percentage of the total loan amount have been forgiven?
  * How did the U.S. House member vote on the legislation initially author zing the PPP loan program.

Feel free to add any interesting "color" (little interesting details) you find in your research.  

Some hints to help you answer this question:

* West Virginia's three U.S. House members are listed [here](https://www.govtrack.us/congress/members/WV#representatives)
* To find businesses they or their families own, you can search for financial disclosure forms for U.S. House members [here](https://disclosures-clerk.house.gov/PublicDisclosure/FinancialDisclosure#Search).
  * To speed things up, here are direct links to the three U.S. House members most recent financial disclosures, which I found by searching the site linked above.
  * [Rep. 1](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10042042.pdf)
  * [Rep. 2](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10040330.pdf)
  * [Rep. 3](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10042734.pdf)
* When you examine the disclosure forms, there may be lots of stock sales and retirement accounts mixed in.  Ignore those. In the table in the disclosure forms, look for "Partnership" or "C-Corp" in the column named "Income Type"
* This is a link to details on passage of the [CARES ACT](https://www.congress.gov/bill/116th-congress/house-bill/748). You can find a link to how the House voted by clicking on a link from [this page](https://www.congress.gov/bill/116th-congress/house-bill/748/actions).

A3: Rep. Carol Devine Miller’s husband Matt Miller is the owner of multiple auto dealerships that received PPP loans, including: DM MOTOR, INC, DUTCH MILLER OF CHARLESTON, INC, DUTCH MILLER OF CHARLOTTE, INC, DUTCH MILLER SUBARU, INC, and DUTCH MILLER CHEVROLET, INC. The five businesses combined to receive $3,105,300 total. In the end, 5 Dutch Miller got 101.13% of its loans forgiven, or in other words made an extra $35,171. On April 23, 2020, Rep. Miller voted in favor of the legislation, and then blamed Congressional Democrats when it didn't pass for 14 days.

```{r}
miller_motors <- wv_ppp_loans %>%
  mutate(new_name = case_when(
    name == "DUTCH MILLER OF CHARLESTON, INC" ~ "Dutch Miller",
    name == "DUTCH MILLER OF CHARLOTTE, INC" ~ "Dutch Miller",
    name == "DUTCH MILLER SUBARU, INC" ~ "Dutch Miller",
    name == "DUTCH MILLER CHEVROLET, INC." ~ "Dutch Miller",
    name == "DM MOTOR, INC" ~ "Dutch Miller",
    TRUE ~ name
  )
  ) %>% 
filter(new_name == "Dutch Miller") %>% 
   # I used summarise(sum(amount) to get the amount total of 3105300.
 # I tried to use summarise(sum(forgiveness_amount)) but I got the following error message: "Error: negative length vectors are not allowed" --> So I added all of the forgiveness amount values in the console and got 3140471.
#I then divided the forgiveness total (3140471) by the total loan amount (3105300) in the console, and multiplied it by 100, which gave me 101.13%
  select(amount, forgiveness_amount) %>%
  mutate(
  percent_forgiven = (forgiveness_amount/amount)*100
)
#and then just to double check, I took the amounts I got in the command above, added them all up in the console, and divided them by 5. Also got 101.13%.


# Hint
# First, create a dataframe that only has loans to those five companies
# Second, use that newly created dataframe to calculate total loan amount and percentage forgiven for those companies

```
