---
title: "lab_05"
author: "Sean Mussenden"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and janitor.

**Task** load these two packages.

```{r}
# Turn off scientific notation
options(scipen=999)
#install.packages('tidyverse')
install.packages('janitor')
library(tidyverse)
```

## Load Data

You'll need to load three data sets for this:

* The West Virginia slice of the PPP loan data (lab_05.rds).
* A "lookup table" that allows you to translate NAICS (industry) numeric codes to industry titles (naics_codes.csv).
* A table of West Virginia population by county (American Community Survey, 2019 5-year averages) (wv_population_county.csv).

All three data sets are in the data folder.  Write code to load the three in the codeblock below.

**Task** Create a codeblock below this task, then read the data sets in in and assign them to appropriate variable names. There's a fourth data set you'll use in this lab, too, of selected loans in ZIP Code 25401. But there's no need to load it now.
```{r}
wv_ppp_loans <- read_rds("data/lab_05.rds")
naics_codes <- read_csv("data/naics_codes.csv")
wv_population_county <- read_csv("data/wv_population_county.csv")
```

## Answer questions

**Q1.** In the data folder, there is a csv called zip_25401_loan_sample.csv.  It contains a sample of loans from West Virginia ZIP Code 25401.

As we read earlier this semester, [multiple loan applications coming from multiple businesses at the same residential street address](https://www.nytimes.com/2021/08/17/business/ppp-fraud-covid.html) might point to fraud. Or it could alert us to companies that used [multiple corporate entities to get more money than envisioned](https://www.propublica.org/article/different-names-same-address-how-big-businesses-got-government-loans-meant-for-small-businesses) by the original law that authorized the program.   

You are going to examine this data to see if we can find a lot of loans coming from the same address.  Here's the problem: the street address field is pretty messy.  The same address appears with minor variations --  "1003 Sushruta Dr" vs "1003 SUSHRUTA DR" -- that will prevent proper grouping.

First, upload the data into Open Refine and standardize/clean the address field - make sure to make a copy of that column to work on and to try all of the options for clustering. If you've done it properly, you should have 65 discrete addresses.

Then export the data from Open Refine, and move it to the proper folder.

Next, load it in the codeblock below, assigning it to an appropriate variable name.

```{r}
zip_new <- read_csv("data/or-zip-25401-loan-sample-csv.csv")
```

Then answer these questions:
* What is the street address in this data that has the most loans? 
* How many loans are there at that street address?
* What are the names of the businesses at that address?

```{r}
zip_new %>%
  group_by(address) %>%
  summarise(
    count_loans = n()) %>%
  arrange(desc(count_loans))
```
```{r}
zip_new %>%
  filter(address == "126 E Burke St")
```


**A1.**
126 E Burke St has the most with 6 loans. The six businesses are:
DREW HOLDINGS LLC
BRIX27, LLC
ABRAHAM ASHTON
HUB CO-OP LLC
HUB CO-OP, LLC
RONIN PROPERTIES LLC


**Q2.** Do some web research on the businesses that answered question 1.  

Google the street address.  Google the business names and search for their corporate records on [Open Corporates](https://opencorporates.com/). Be sure to find the website of the name of the company that appears twice in the list of businesses and develop an understanding of what it does.

Based on your research, does it seem suspicious that this collection of businesses all got loans using the same address? Why or why not. Use specific facts identified in your research to support your position.

**A2.** The building is listed on Google as the Hub Co-Op building, basically a localized WeWork. That to me would immediately signal that multiple businesses are located within the space, so I wouldn't think its too strange that multiple applications came from the same address. Drew Holdings LLC is a lessor of real estate, again not that strange if they deal in renting out property to other businesses. It is a bit strange that Abraham Ashton has two offices listed on the public records, neither of them are the address of the Co-Op. It's enough to look into but I wouldn't think its too much of a red flag.

**Q3.** Start by using the West Virginia slice of the PPP loan data that you loaded at the start of the lab to create a subset of PPP loans in West Virginia's second largest county (which you can find in wv_population_county). And then use that table you created to answer the following questions:

* Which city in that county had the highest number of loans?
* In that city, which industry title had more loans than any other industry title?

Requirement: you MUST export a dataframe of PPP loans from R Studio at some point in the process (not necessarily at the beginning!), load it into Open Refine, clean the city column, export it from Open Refine, and reimport into R Studio. To export data from R Studio, you will use the write_csv() function, which works like this if we're using the wv_population_county dataframe:

Guidance: there are a lot of steps you'll need to take to answer this question. You may find it helpful to write out in English what you plan to do step-by-step before you start writing code.   

```{r}
wv_ppp_with_naics <- wv_ppp_loans %>% left_join(naics_codes, by="naics_code")
#joining www_ppp_loans with naics_codes
```

```{r}
berkeley_loans <- wv_ppp_with_naics %>%
  filter(project_county_name == "BERKELEY")
#creating a new dataset with just loans that appear in Berkeley County, listed in wv_population_county as having the second highest amount of loans.
```

```{r}
# write_csv accepts 2 arguments: the dataframe and the path to the file you write out
write_csv(berkeley_loans, "data\berkeley_loans_csv.csv")
#exporting the berkeley_loans to a csv to clean in open refine.
```

```{r}
cl_berkeley_loans <- read_csv("data/updated-berkeley-loans-csv.csv")
#importing the new cleaned berkeley loans after cleaning them in open refine.
```

```{r}
cl_berkeley_loans %>%
  group_by(city) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))
#grouping the cleaned berkeley loans by city, couting the amount of loans in each city, then arranging them in descending order of number of loans that appear in each city.
```

```{r}
cl_berkeley_loans %>%
  filter(city == 'Martinsburg') %>% 
group_by(title) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))
#filtering the berkeley loans in just Martinsburg, grouping by industry type (title), then summarizing them by counting the number of loans in each industry and arranging them in descending order of number of loans in each industry.
```




**A3.** Within Berkeley County, Martinsburg had the highest number of loans with 1166. Full-Service Restaurants had the most loans for any business type (naics_code) with 50. 

**Q4.** What are your two best hypotheses, which you'd need to confirm with reporting and further analysis, that explain why that industry is at the top of the list?

**A4.** My first hypothesis is that restaurants suffered the most when the first wave of lockdowns hit. On 3/17/2020, WV Gov. Justice announced the closure of all restaurants and bars. A day later, he requested WV ABCA temporarily suspends regulatory rule prohibiting bars and restaurants from providing beer and wine for off-premises consumption, meaning they were looking for ways to keep businesses open while being shut down. It wasn't until over a month later on 4/28/2020 that he provided additional guidelines for the reopening of restaurants with takeaway service or outdoor dining options. My second hypothesis is that, as was happening throughout the country, restaurant workers were some of the highest risk workers and therefore spent a high amount of time away from work, meaning restaurants were short-staffed even when they re-opened.

**Q5.** Start with a dataframe of loans to all businesses in the city and industry that answered question 3. Answer the following questions:

```{r}
restaurants_in_martinsburg <- cl_berkeley_loans %>%
filter(title == 'Full-Service Restaurants') %>%
filter(city == 'Martinsburg')
```

```{r}
restaurants_in_martinsburg %>%
    arrange(desc(amount))
#arrange the loan amounts in descending order of amounts within the restaurants in martinsburg dataset.
```

* What is the name of the business that got the highest approved loan amount? - answered
* How much was it for? - answered
* When was it approved? - answered
* How many jobs does the data say were retained? - answered
* Is there a difference between the business' name in the PPP data and the name its customers know it by? If so, what is that name? - answered
* How many locations does this business have? - answered
* Did one of its locations close during the pandemic, either before or after it got the loan? - answered

Hint: you will not find the answers to the last three questions in the data.  You could call them directly to get that information, but I don't want you to do that for this assignment.  Instead, do some web research. I would start by Googling the company name from the data and looking at the page that comes up for the business from at http://apps.sos.wv.gov/. I would use information I found on that page and use info about the company from Google, the [Wayback machine](https://archive.org/web/) (which lets you look at older versions of a company's website), Yelp, and Facebook.

**A5.** COTTLE CHRISTI L LLC got the highest loan in Martinsburg for 280,434.00. It was approved on February 17th, 2021. The data say there were 94	jobs retained. It looks like it is known by various versions of the name KITZIE'S. There are 6 business locations listed under this LLC - KITZIE'S CAFE, KITZIE'S CAFE' II, KITZIE'S OF INWOOD, KITZIE'S OF SPRING MILLS, KITZIE'S RESTUARANT & LOUNGE, and RIVERBEND BAR & GRILL. It looks like KITZIE'S CAFE' II closed before the pandemic in 2018, and KITZIE'S OF INWOOD closed during the pandemic on 5/19/2020. 
