---
title: "lab_10.Rmd"
author: "derek willis"
date: "11/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Datawrapper account
* PPP data that is in a CSV file in the `data` folder
* Other data in CSV files in the `data` folder

## You might need

* A Census API key

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
library(tidyverse)
library(tigris)
library(tidycensus)
library(lubridate)
```

```{r echo=FALSE}
census_api_key("ab0ed8418e6554d3eb77ed342784e3bb0b002a64")
#shouldn't need to do run this twice, don't need to do install=TRUE the second time.
#if you get a note saying its stored but need to restart R, copy readRenviron("~/.Renviron") and run it in the console. 
```

```{r}
all_counties <- read_csv("data/all_counties.csv")
barber_shops <- read_csv("data/barber_shops_ppp.csv")
county_fips <- read_csv("data/county_fips.csv")
kabbage <- read_csv("data/kabbage_counties.csv")
```

## Questions

**Q1.** You got a tip that there are some unusual patterns among applicants listing barber shops as their occupation. Which state has the most PPP applications from barber shops, and how many of those have undisbursed amounts greater than $0? Which state has the most undisbursed applications?

You MUST do the following things when answering this question:
* Use the PPP loan data, which is in the data folder and is called `barber_shops_ppp.csv`.
* Using the barber shop PPP applications, write code that generates two dataframes:
  * one with a total number of applications per state
  * one with a total number of applications per state where the undisbursed amount is more than 0.
  
```{r}
#one with a total number of applications per state
grouped_barber_shops_by_state <- barber_shops %>% 
  group_by(state) %>%
  summarise(
    total_loans = n()
  ) %>% 
  arrange(desc(total_loans))
```

```{r}
#one with a total number of applications per state where the undisbursed amount is more than 0.
undispersed_barber_shops <- barber_shops %>% 
  filter(undisbursed_amount > 0) %>%
  #total loans = 11827
  group_by(state) %>%
  summarise(
    total_loans = n()
  ) %>% 
  arrange(desc(total_loans))
```  
  
* Join those dataframes, specifying the join column, and rename the totals columns so that they better describe the values in them.
```{r}
#then join them 
joined_und_by_state_and_total_bshops <- grouped_barber_shops_by_state %>% 
  left_join(undispersed_barber_shops, by=c('state')) %>% 
  rename(total_nloans_per_state = total_loans.x) %>%
  rename(nloans_outstanding_und = total_loans.y)
```

* Export the top 10 states (in terms of total number of applications) to a CSV file using `write_csv`

```{r}
joined_und_by_state_and_total_bshops %>%
  #head(10)
  filter(total_nloans_per_state >= 2448)

joined_und_by_state_and_total_bshops
```
```{r}
write_csv(joined_und_by_state_and_total_bshops, "data/bshop_loans_and_undispersed_amounts_by_state.csv")
```

* Using the top 10 states, make a bar chart with Datawrapper. What kind of bar chart is up to you, so experiment with a few kinds of bar charts to see what answers the question above in the clearest way. Your chart MUST:
  * Have a title, description and source that tells readers about the data. The title and description should highlight the most newsworthy or important aspect.
  * Easily readable labels and values
  * Enough spacing so that the bars are visually distinct
  * Use color for contrast and make it clear what each color represents
* Publish your chart (you may need to confirm your Datawrapper account if you haven't already in order to do this)
* Put the answer below along with the URL of your chart.


**A1.** Illinois has the most PPP applications from barber shops with 11,080. Illinois also has the highest amount of undispersed loans with 1,713. Of all of the barbershop loans nationwide, 11,827 loans have undispersed amounts greater than 0.
 
Q1 chart - https://www.datawrapper.de/_/PBbqg/ (Unfortunately despite my best efforts, I couldn't manually change the name of the label key from the original column names to make it more readable!)


**Q2.** Using Datawrapper, make a different (ideally better) version of the chart you made above. TO BE CLEAR: this new graphic should show information about states and loans with undisbursed amounts, so don't, for example, focus on a single state. To do this, you MUST go back in R and do any calculations or add additional data to make the chart more effective or answer the above question in a different way. Your chart MUST be different in that it cannot be a bar chart showing the counts of total and undisbursed applications by state. It could be a bar chart showing another calculation or metric, or a scatterplot, or some other form. Follow the same general guidelines as above for the chart's visual elements, and publish it and including the URL below, along with an explanation of how this chart is different from the first one and how you decided what to do.
```{r}
percents <- joined_und_by_state_and_total_bshops %>% 
mutate(percentages = (nloans_outstanding_und/total_nloans_per_state)*100
) %>% 
  filter(total_nloans_per_state > 10) %>% 
  filter(percentages > 15) %>% 
  select(state, percentages)
```

```{r}
write_csv(percents, "data/percentages_of_undispersed.csv")
```


**A2.** https://www.datawrapper.de/_/bj9Jm/

I went with a scatter plot that had the numbers displayed next to the point to show the percent that undispersed barbershop loans accounted for the overall number of barbershop loans per state. 


**Q3.** The online lender Kabbage, Inc. is featured in the [ProPublica PPP fake farms story](https://www.propublica.org/article/ppp-farms). To show where Kabbage was lending money, make a county-level map that shows the percentage of loans in each county that were Kabbage loans. To do this, make a codeblock and do the following:

1. Load the `kabbage_counties.csv` and `all_counties.csv` files in the data folder.
(done, above)

2. Join the two together, specifying the join columns, and rename the total columns to represent Kabbage loans and total loans. Save the output to a new dataframe.

```{r}
nkabbageloans_and_ncountyloans <- all_counties %>% 
  left_join(kabbage, by=c('project_county_name', 'project_state')) %>% 
  
  rename(kabbage_loans_per_county = total.y) %>%
  rename(all_loans_per_county = total.x)
```

3. Calculate the percentage of loans in each county that were Kabbage loans and save the output to the existing dataframe.

```{r}
nkabbageloans_and_ncountyloans <- nkabbageloans_and_ncountyloans %>% 
  mutate(
  perc_kabbage_loans_per_county = (kabbage_loans_per_county/all_loans_per_county)*100
)

```

4. Load the `county_fips.csv` file in the data folder into a dataframe. (done)

5. Join the combined counties data with the fips data, again specifying the join columns and saving it to a new dataframe.

```{r}
county_fips <- county_fips  %>%
  mutate(match_county = str_replace(match_county, ' COUNTY', '')) %>%
  #replaced "COUNTY" with a blank space aka nothing
   mutate(match_county = str_replace(match_county, "'", "")) %>%
  #took out all apostrophes, like "Queen Anne's"
   mutate(match_county = str_replace(match_county, "^ST. ", "SAINT "))
#spelling out ST. in Saint Mary
```


```{r}
#for if we had to read them in from tigris
#all_fips <- fips_codes %>% as_tibble() %>%
  #mutate(fips_code = str_c(state_code, county_code)) %>%
  #mutate(match_county = str_to_upper(county)) %>% 
  #mutate(match_county = str_replace(match_county, ' COUNTY', '')) %>%
  #replaced "COUNTY" with a blank space aka nothing
  # mutate(match_county = str_replace(match_county, "'", "")) %>%
  #took out all apostrophes, like "Queen Anne's"
   #mutate(match_county = str_replace(match_county, "^ST. ", "SAINT "))
#spelling out ST. in Saint Mary
```

```{r}
fips_with_loans <- county_fips %>% 
  left_join(nkabbageloans_and_ncountyloans, by=c('match_county'='project_county_name', 'state'='project_state')) %>% 
  select(fips_code, match_county, perc_kabbage_loans_per_county)
```


6. Export the new dataframe to a CSV file.
```{r}
write_csv(fips_with_loans, "data/fips_with_kabbage_loans_v2.csv")
```

7. Make a county-level map using that exported CSV file that displays the percentage of Kabbage loans. Style it so that the tooltip displays the county and the percentage, and format all percentages shown on the map so that they appear with one decimal point (such as 3.4%). Follow the same general guidelines as above for the chart's visual elements, and give the map a title, description and source. Provide a legend caption and format the percentage in the same way as on the tooltips.


8. In the answer space below, provide the URL for your published map and then describe where Kabbage loans are and are not found. Where are they more common? Look specifically at Florida in the data: are there counties on the map where no data is displayed? Why?

**A3.**  https://www.datawrapper.de/_/xAzx4/  Kabbage loans are definitely more common around Atlanta and around Miami more than anywhere else. Kabbage is an Atlanta-based company, so they may tend to offer loans to businesses in the Atlanta area, which spills out into neighboring South Carolina and Alabama. The peninsula of Florida seems to have more loans than the panhandle, assuming because they have a greater concentration of orange farms, since other major cities don't have the same consistency of numbers that Atlanta and Miami have. Interestingly, suburban New Jersey also has a lot, but it looks to be separated by the state boundary of New Jersey. It doesn't really make sense that a county in New Jersey across from Philly would be drastically different than a county in Pennsylvania right near Philly. Looking also at cities, Los Angeles doesn't have too many loans and you can see the outline of suburban Chicago, which also doesn't have that many loans. The loans mostly follow the eastern seaboard, up the length of the Appalachian Mountains. Since Kabbage deals in farming loans, I'm guessing most of the applications were dealing with farming along the mountain range, plus the orange groves in Florida. 
